---
title: "Tables and Figures - Supplement A and B"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


This document contains the code to reproduce tables and figures from Supplement A and Supplement B for the project "Heterogeneity in direct replications in psychology and its association with effect size". 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
Sys.setlocale("LC_TIME", "English" ) #to print date in YAML in English
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kablExxtra")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(cowplot)){install.packages("cowplot")}
if(!require(boot)){install.packages("boot")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(ggplot2) #plotting figures
library(cowplot) #combine plots
library(kableExtra) #create fancy pdf tables
library(boot) #Bootstrap confidence intervals for correlations

dat <- read_csv("../data/collated_summary_data.csv")

source("03_helper_functions_tables_figures.r") #Load functions to prep data for figures
```

## Supplement A

```{r a-plot-prep, cache = TRUE}
res <- dat %>%
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

het <- dat %>%
  select(rp, Site, effect, effect_type) %>%
  group_by(rp, effect) %>%
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>%
  left_join(res) %>%
  mutate(effect_type = recode(effect_type,
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>%
  ungroup()

a <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")
```


```{r, cache = FALSE}
a_plot <- ggplot(a, aes(x = s_I2, y = sqrt(tau2))) + geom_point() + ggtitle("SMD (untransformed)") +
  scale_x_continuous(name = expression(paste(italic(I)^2, "Index"))) +
  scale_y_continuous(name = "Tau")

```


```{r trunc-biserial, cache = TRUE}
#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>%
  filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales')) %>%
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")



#Add rp and k to the data-frame
het <- dat %>%
  select(rp, Site, effect, effect_type) %>%
  group_by(rp, effect) %>%
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>%
  left_join(res2) %>%
  mutate(effect_type = recode(effect_type,
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>%
  ungroup()

d <- het %>% filter(effect_type == "SMD.") %>%
  mutate(sigma2 = tau2/(I2/100) - tau2) #add 'typical within study variance' per project

```


```{r c-plot-1, cache = FALSE}
c_plot <- ggplot(d, aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("Point Biserial r (transformed SMD)")

```


```{r tau-vs-tau}
b <- select(a, effect, tau2) %>%
  left_join(select(d, effect, tau2), by = "effect") #add truncated transformation tau2
```


```{r b-plot-1}
b_plot <- ggplot(b, aes(x = sqrt(tau2.x), y = sqrt(tau2.y))) +
  geom_point() +
  ggtitle("SMD vs. Point biserial") +
  scale_x_continuous(name = "Tau (SMD)") +
  scale_y_continuous(name = "Tau (Point biserial)")

```




```{r, fig.width = 6, fig.asp = 1.3}
cowplot::plot_grid(a_plot, b_plot, c_plot, ncol = 1, labels = "AUTO")


# ggsave("../figures/Figure-S1.png", dpi = 600, width = 6, height = 7.8)
```

_Figure S.1_ Association between estimates of $\tau$ and $I^2$ after and before standardized mean differences (SMD) effects were transformed into biserial correlations for all effects reported as SMDs in ManyLabs 1. Panel A shows the association between estimates of $\tau$ and $I^2$ without transformation when using the meta-analytic specification of the original authors (see main text). Panel B shows the association between estimates of $\tau$ for the same effects before and after transformation to biserial correlations, either with truncation of correlations over 1 (circles) or without (crosses) when estimating variance. Panel C  shows the associations between estimates of $I^2$ on the SMD metric and estimates of $\tau$ after transformation, either with truncation or not.



#Table S1.
###Heterogeneity across primary effects and statistical power of ten multi-lab replication projects, ordered with respect to estimated heterogeneity ($I^2$).
```{r prep-heterogeneity-table, cache = TRUE}
#library(metafor)
#library(dplyr)
#library(purrr)

#fit a random effects model to all empirical data
#Each set of input datapoints is meta-analyzed as per the original replication project
#Function 'est_heterogen_smd_raw' is sourced
res <- dat %>%
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") %>% #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date
  mutate(tau = sqrt(tau2),
         tau_ci.lb = sqrt(tau2_ci.lb),
         tau_ci.ub = sqrt(tau2_ci.ub)) %>%
  select(-tau2, - tau2_ci.lb, -tau2_ci.ub) #drop redundant variables

#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>%
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")

res <- res %>%
  mutate(r = res2$r,
         tau_r = sqrt(res2$tau2), #For table 3, report tau and transformed effect sizes (biserial correlations)
         tau_r_ci.lb = sqrt(res2$tau2.lb),
         tau_r_ci.ub = sqrt(res2$tau2.ub))

#Add rp and k to the data-frame and prep for table
het <- dat %>%
  select(rp, Site, effect, effect_type) %>%
  group_by(rp, effect) %>%
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>%
  left_join(res) %>%
  mutate(effect_type = recode(effect_type,
                              d = "SMD",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>%
  ungroup()

#Load simulation results (power/type 1 error for all effects)
dens <- readRDS("../data/power_simulation_results.RDS")
names(dens) <-  sort(unique(dat$effect)) #names were alphabetized upon splitting in simulation so must sort

#Summary of simulation results, needs to be combined with main part of table
I2_power <- dens %>%
  bind_rows(.id = "effect") %>%
  group_by(effect, tau_index) %>%
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>%
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(zero = '1', small = '2', medium = '3', large = '4')

#Combine dataframes to create full table
het <-  het %>%
  left_join(I2_power) %>%
  arrange(desc(s_I2),  desc(s_ci.ub)) %>% #sort by I2 and upper CI bound
  mutate(I2_ci = paste0("[", trimws(format(round(s_ci.lb, 1), nsmall = 1)), ", ", trimws(format(round(s_ci.ub, 1), nsmall = 1)), "]"), #Round and put confidence interval between brackets, e.g., [0, 0.41]
         s_I2 = format(round(s_I2, 1), nsmall = 1),
         tau_ci = paste0("[", trimws(format(round(tau_ci.lb, 3), nsmall = 3)), ", ", trimws(format(round(tau_ci.ub, 3), nsmall = 3)), "]"),
         tau = format(round(tau, 3), nsmall = 3),
         tau_r = format(round(tau_r, 3), nsmall = 3),
         tau_r_ci = paste0("[", trimws(format(round(tau_r_ci.lb, 3), nsmall = 3)), ", ", trimws(format(round(tau_r_ci.ub, 3), nsmall = 3)), "]")) %>% #rounded here instead of in table function because should have 4 digits
  select(rp, effect, k, effect_type, eff_size, s_I2, I2_ci, tau, tau_ci, r, tau_r, tau_r_ci, zero, small, medium, large)

```


```{r heterogeneity-table, cache = FALSE}
#library(kableExtra)

het %>% knitr::kable(align = "c", booktabs = TRUE, escape = FALSE, digits = 2, col.names = c("RP", "Effect", "K", "Effect type", "Effect size estimate",  "$I^2$(\\%)", "$I^2$ 95\\% CI", "$\\tau$", "$\\tau$ 95\\% CI", "r*" , "r* $\\tau$", "r* $\\tau$ 95\\% CI", "Zero", "Small", "Medium", "Large")) %>%
  add_header_above(c(" " = 12, "Level of heterogeneity" = 4)) %>%
  add_header_above(c(" " = 12, "Type I Error Rate & Statistical Power" = 4)) %>%
  kable_styling("scale_down", font_size = 7) %>% #comment out if outputting html file
  column_spec(5, width = "5em") %>% #make sure the 'Effect size estimate' header is wrapped
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, K = no. primary studies, CI = confidence intervals, r* = effect sizes as correlations and biserial correlations. $I^2$ and its confidence intervals are for original effect size specifications whereas $\tau$ and its confidence intervals are for transformed r* effect sizes. Statistical power and type I error rates were simulated, where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75\\\\%) respectively. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Code to reproduce table: https://osf.io/j9zyd/?view\\\\_only=e6639d8d4f924739a7310782fbbb4e06",
           alphabet = "These effects were simulated as standardized mean differences") %>%
  landscape()


```

## Supplement B


```{r Supplement-B-table}
#Load simulation results (power/type 1 error for all effects)
dens <- readRDS("../data/power_simulation_results.RDS")
names(dens) <-  sort(unique(dat$effect)) #names were alphabetized upon splitting in simulation so must sort


#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4', Effect = effect)

##Medium effect size results
dens2 <- readRDS("../data/supplementB_power_simulation_results.RDS") 
names(dens2) <-  sort(unique(dat$effect))

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4') %>% 
  select(-effect)

##combine for table

comparison <- cbind(zero, medium)
```

Sensitivity analysis assuming a medium true effect size rather than zero true effect size

In our main results we simulated data to estimate type I error and power using an average true effect size equal to zero. As explained in the methods section this is unlikely to matter as our interest lies in heterogeneity. However, as a sensitivity analysis, in Supplement B we also estimated type I error and power, but generated with an average 'medium' true effect size as defined by Cohen (1988).

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table SB1. In both cases type I error is approximately nominal, as compared to the expected 5% error rate. With a true effect size of zero, average power was `r 100*round(mean(zero$Small), digits = 2)`%/`r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`% for small/medium/large heterogeneity, respectively. The corresponding power with a true medium effect size was also `r 100*round(mean(zero$Small), digits = 2)`%/ `r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`%, respectively. To conclude, changing true effect size in our simulations from zero to medium had little to no impact on estimated type I error and power.

_Type I error and power to detect four heterogeneity levels under two true effect sizes_

```{r}
knitr::kable(comparison, digits = 2)
```

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Medium effect size is as defined by Cohen (1988). Code to reproduce table: https://osf.io/j9zyd/?view_only=e6639d8d4f924739a7310782fbbb4e06

## Supplement C

```{r, preprocessing}

SMDs_raw <- dat %>% filter(outcomes1_2 == "mean _ SD") %>% 
  mutate(wi = compute_precision(SD1 = outcome_t2, SD2 = outcome_c2, n1 = ntreatment, n2 = ncontrol)) #compute_precision function is sourced

average_within <- SMDs_raw %>% #these values we now wish to correlate with meta-analytic results
  group_by(effect) %>% 
  summarize(s2 = sum(wi) * (n()-1) / (sum(wi)^2 - sum(wi^2))) %>% #Average within study variance; equation 9. Higgins and Thompson (2003). n() = k [number of studies in a meta-analysis]
  ungroup()


#Raw mean differences to SMD
MD_standardized <- dat %>% 
  filter(outcomes1_2 == "mean _ SD") %>% #Note, drops Mean _ SE studies (two) since no method in metafor to convert into standardized mean differences
  split(.$effect) %>% 
  map_dfr(MD_fit, .id = "effect") #MD_fit function is sourced

tau_per_type <- MD_standardized %>% 
  left_join(average_within) %>%
  mutate(sigma = sqrt(s2),
         eff_size = abs(eff_size),
         tau = sqrt(tau2)) #needed for bootfitter function below to work but not used

#Add labels
#lables
set.seed(499) #for the bootstrapping
all_effects_ci <- tau_per_type %>% 
  bootfitter_sigma(.) #bootfitter is a sourced function

labels_all <- tau_per_type %>% #
  fitter_sigma(.) %>% #fitter is a sourced function
  mutate(lab = paste0("r = ", r, " ", all_effects_ci$ci)) #put output into one variable for improved plotting


#As above but dropping the 4 largest effect sizes (Anchoring effects)----
without_anchoring <- tau_per_type %>% 
  filter(!grepl("Anchoring", effect))

#lables
set.seed(500) #for the bootstrapping
restricted_effects_ci <- without_anchoring %>% 
  bootfitter_sigma(.) #bootfitter is a sourced function

labels_restricted <- without_anchoring %>% #row 1:2 = tau2, row 3:4 = I2, row 5:6 = H2
  fitter_sigma(.) %>% #fitter is a sourced function
  mutate(lab = paste0("r = ", r, " ", restricted_effects_ci$ci)) #put output into one variable for improved plotting


```

```{r}

within_plot_full <- ggplot(tau_per_type, aes(x = eff_size, y = sigma)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") + #to set alpha for the line geom_line must be used rather than geom_smooth
  geom_text(data = labels_all[labels_all$index == "sigma",], aes(label = lab), x = -Inf, y = Inf, hjust = -0.05, vjust = 2, size = 3.5, fontface = "italic", alpha = .6) + #hjust and vjust can be used to annotate the same position in all facets even though the scale are different
  scale_y_continuous(name = expression(paste("Within Studies Standard Deviation"))) +
  scale_x_continuous(name = "Absolute Effect Size") +
  theme_classic() +
  theme(strip.background = element_blank(), #remove box around titles
        strip.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, colour = "black", #add box around plots
                                    size = 0.5, linetype = "solid"))


without_anchoring_plot <- ggplot(without_anchoring, aes(x = eff_size, y = sigma)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") + #to set alpha for the line geom_line must be used rather than geom_smooth
  geom_text(data = labels_restricted[labels_restricted$index == "sigma",], aes(label = lab), x = -Inf, y = Inf, hjust = -0.05, vjust = 2, size = 3.5, fontface = "italic", alpha = .6) + #hjust and vjust can be used to annotate the same position in all facets even though the scale are different
  scale_y_continuous(name = expression(paste("Within Studies Standard Deviation"))) +
  scale_x_continuous(name = "Absolute Effect Size") +
  theme_classic() +
  theme(strip.background = element_blank(), #remove box around titles
        strip.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, colour = "black", #add box around plots
                                    size = 0.5, linetype = "solid"))

cowplot::plot_grid(within_plot_full, without_anchoring_plot, nrow = 2, labels = "AUTO")
ggsave("../figures/figure-S2.png", dpi = 600, height =  7, width = 7)
```

Figure S2. The Pearson correlation (_r_) between average absolute effect size and average within studies standard deviation for 40 effects that were measured as mean or standardized mean differences. Panel A shows results for all 40 effect sizes and panel B with the four largest (Anchoring 1 - 4) excluded. Effects reported as mean differences were standardized. Square brackets contain 95% bootstrapped percentile confidence intervals. 