---
title: "Troubleshooting tau2"
author: "Anton Olsson Collentine"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}
if(!require(writexl)){install.packages("writexl")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables
library(ggplot2)
library(cowplot)
library(writexl)


dat <- read_csv("../data/collated_summary_data.csv")

source("helper_functions_tables_figures.r") #Load functions to prep data for figures
```




```{r, cache = TRUE}
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

a <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.") 

a_plot <- ggplot(a, aes(x = s_I2, y = sqrt(tau2))) + geom_point() + ggtitle("ML1 SMD - untransformed effect sizes")

```


```{r prepping, cache = TRUE}

#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")



#Add rp and k to the data-frame 
het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res2) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

b <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.") 

b_plot <- ggplot(b, aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("ML1 biserial r - transformed SMD effect sizes")



```

```{r untruncated-biserial, cache = TRUE}


res3 <- dat %>%
  filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & outcomes1_2 == "mean _ SD") %>% 
  split(.$effect) %>% 
  map(~transform_SD(.$outcome_t1, .$outcome_c1, .$outcome_t2, .$outcome_c2, .$ntreatment, .$ncontrol)) %>% 
  map(~untruncated_d_to_r(.$d, .$n1, .$n2)) %>% 
  map_dfr(summarizer, .id = "effect")

c_plot <- b_plot + geom_point(data = res3, shape = 4, aes(x = I2, y = sqrt(tau2)))




```

Figure 1.
```{r}
cowplot::plot_grid(a_plot, c_plot, ncol = 1)
```

**The four effects at the end are Anchoring 1,2,3,4. X corresponds to untruncated and circles are truncated**

We see that the non-monotonicity after transformation is not explained by truncation. Rather (especially in view of the 4th to last effect which wasn't truncated) there seems like there might be a relation with effect size.

##Effect size?
```{r, cache = TRUE}

both <- dat %>%
  filter(effect %in% c("Quote Attribution", "Anchoring 1 - NYC")) %>% 
  split(.$effect) %>% 
  map(function(x){
    metafor::escalc(measure = "SMD", m1i = x$outcome_t1, m2i = x$outcome_c1, sd1i = x$outcome_t2, sd2i = x$outcome_c2, n1i = x$ntreatment, n2i = x$ncontrol, data = x)
    })

both_trunc <- both %>% 
  map(~transform_d_to_r(.$yi, .$ntreatment, .$ncontrol)) 

both_nontrunc <- both %>% 
  map(~untruncated_d_to_r(.$yi, .$ntreatment, .$ncontrol)) 



#For excel
out <- bind_cols(both[[1]][, c("yi", "vi")], both[[2]][, c("yi", "vi")],
          both_trunc[[1]][, c("r", "vi")], both_trunc[[2]][, c("r", "vi")],
          both_nontrunc[[1]][, c("r", "vi")], both_nontrunc[[2]][, c("r", "vi")],
          both[[1]][, c("ntreatment", "ncontrol")], both[[2]][,c("ntreatment", "ncontrol")]) %>% 
  mutate_at(vars(contains("vi")), sqrt) %>% 
  select(yi, vi, yi1, vi1, r, r1, everything(), -r2, -r3) %>% 
  rename("Anch_d" = yi, "Anch_d_SE" = vi, "Quote_d" = yi1, "Quote_d_SE" = vi1,
         "Anch_r" = r, "Anch_trunc_SE" = vi2, "Quote_r" = r1, "Quote_trunc_SE" = vi3,
         "Anch_nontrunc_SE" = vi4, "Quote_nontrunc_SE" = vi5,
         "Anch_n1" = ntreatment, "Anch_n2" = ncontrol, "Quote_n1" = ntreatment1, "Quote_n2" = ncontrol1)
  



write_xlsx(out, "effect_sizes_with_SE.xlsx")
```

```{r}
eff <- out %>% 
  mutate(ratio_SE_d = Anch_d_SE / Quote_d_SE,
         ratio_SE_biserial = Anch_trunc_SE / Quote_trunc_SE,
         Quote_ratio = Quote_d_SE / Quote_trunc_SE,
         Anch_ratio = Anch_d_SE / Anch_trunc_SE,
         diff = ratio_SE_d - ratio_SE_biserial) %>%
  arrange(desc(Anch_d)) %>%
  select(Anch_d, Quote_d, Quote_ratio, Anch_ratio, ratio_SE_d, ratio_SE_biserial, diff)

```

- The transformation is not monotonic for large values and this is not explained by the truncation (see next section) which almost only affects the I2

- Let's take a closer look at two effects with very different effect sizes (mean SMD `r round(mean(eff$Quote_d), 2)` vs. `r round(mean(eff$Anch_d), 2)`): Quote attribution [fifth effect from the right Figure 1] vs. Anchoring 1. [4th effect from the right Figure 1].

**Consider the following:**

a) higher variance for large effect sizes -> less weight for these effects -> higher tau2 estimate

b) The formulas for variance for hedge's g and for variance for biserial correlations both include effect size

c) For hedge's g larger effect size -> more variance (see Borenstein for formula) and not clear for biserial

d) If the impact of effect size on variance is larger for hedge's g than for biserial it makes sense to expect some differences in the variance (and tau2) estimates after transformation

e) If the weight of large effect sizes (inverse variance) is relatively higher after transformation (i.e., their relative variance decreases) then we would expect lower tau2 estimates after transformation. This effect should be larger for larger effect sizes. 

f) This is exactly what we see below and in figure 1. In the table below, note the large difference in SE ratios before vs. after transformation to biserial correlations. The relative variance of the larger effect sizes (anchoring) decreases considerably after transformation. The same is true of larger effect sizes _within_ the Anchoring effect: the correlation between the Anchoring effect sizes and the difference in ratios is _r_ = `r cor(eff$Anch_d, eff$diff)`. 

```{r}

knitr::kable(eff, booktabs = TRUE) 
```
Note: Anch_d = SMD Anchoring, Quote_d = SMD Quote attribution, ratio_SE_d = Ratio SMD SEs: Anchoring / Quote attribution, ratio_SE_biserial = Ratio biserial SEs: Anchoring / Quote attribution, diff = Ratio SMD - Ratio biserial.

Correlation Anch_d and Anch_ratio _r_ = `r cor(eff$Anch_ratio, eff$Anch_d)`

##Truncation problems?
**Compare the actual values**
```{r, cache = TRUE}
b %>% select(effect, r, tau2, I2) %>% left_join(res3 %>% select(effect, r, tau2, I2), by = "effect") %>% 
  rename(trunc_ES = r.x, trunc_tau = tau2.x, trunc_I2 = I2.x, nontrunc_ES  = r.y, nontrunc_tau = tau2.y, nontrunc_I2 = I2.y) %>% 
  mutate(trunc_tau = sqrt(trunc_tau), nontrunc_tau = sqrt(nontrunc_tau)) %>% 
  knitr::kable(booktabs = TRUE)


```

It is strange that there are (very slightly) different outcomes for non-truncated effects, this is probably due to metafor (which I use for truncation) uses the exact variance formula whereas I use the approximate. Compare truncated vs. non-truncated manual formulas to check:
```{r}
res4 <- dat %>%
  filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & outcomes1_2 == "mean _ SD") %>% 
  split(.$effect) %>% 
  map(~transform_SD(.$outcome_t1, .$outcome_c1, .$outcome_t2, .$outcome_c2, .$ntreatment, .$ncontrol)) %>% 
  map(~transform_d_to_r(.$d, .$n1, .$n2)) %>% 
  map_dfr(summarizer, .id = "effect")

res4 %>% select(effect, r, tau2, I2) %>% left_join(res3 %>% select(effect, r, tau2, I2), by = "effect") %>% 
  rename(trunc_ES = r.x, trunc_tau = tau2.x, trunc_I2 = I2.x, nontrunc_ES  = r.y, nontrunc_tau = tau2.y, nontrunc_I2 = I2.y) %>% 
  mutate(trunc_tau = sqrt(trunc_tau), nontrunc_tau = sqrt(nontrunc_tau)) %>% 
  knitr::kable(booktabs = TRUE) 
```
Exactly the same result for non-truncated effects, so that seems to be the case

Let's take a closer look at the truncated vs. untruncated variance estimates for Anchoring 4 which has the largest change in I2.

```{r, cache = TRUE}
anch4 <- dat %>%
  filter(effect == "Anchoring 4 - Babies") %>% 
  rowwise() %>% 
  do(transform_SD(.$outcome_t1, .$outcome_c1, .$outcome_t2, .$outcome_c2, .$ntreatment, .$ncontrol)) 
#See this stackoverflow question to understand the usage of rowwise and 'do'
#https://stackoverflow.com/questions/22240816/dplyr-summarise-with-multiple-return-values-from-a-single-function

anch4_trunc <- anch4 %>% 
  do(transform_d_to_r(.$d, .$n1, .$n2))

anch4_nontrunc <- anch4 %>% 
  do(untruncated_d_to_r(.$d, .$n1, .$n2))

bind_cols(anch4_trunc, anch4_nontrunc) %>% 
  rename(trunc_vi = vi, nontrunc_vi = vi1) %>% 
  select(r, trunc_vi, nontrunc_vi, n) %>% 
  arrange(desc(r)) %>% 
  knitr::kable(booktabs = TRUE)


```
Note that the (biserial) r itself is not truncated and so the same between the two. It is clear truncation leads to a larger estimate of the variance for an effect. 

**To summarize: Truncation leads to an increase in estimated study variance, but a (small) decrease in meta-analytic effect size estimate. In other words, truncation (as seen in the first figure) leads to a lower I2 but does not affect the $\tau$ particularly**



##Old stuff

Check which effects get truncated:

```{r, echo = TRUE, cache = TRUE}
dat %>% 
  split(.$effect) %>% 
  map_dfr(transform_MA, .id = "effect") %>% 
  filter(r > 1)
```

Basically the anchoring effects.


```{r, }
het %>% filter(rp == "ML3") %>% ggplot(aes(x = I2, y = sqrt(tau2))) + geom_point() + facet_wrap(~effect_type) + ggtitle("ML3 (SMD transformed to biserial correlations)")
```

**The one non-increasing effect in the left plot above is 'Credentials interaction'. **
However, note that all the ML3 'correlations' are either transformed partial eta-squared of 2x2 tables in the first table (we can't fix this, would have to get into the raw data which seems liek too much work, I mentioned this in the revised draft)



```{r}
het %>% ggplot(aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("All effects - tetrachoric, biserial & product-moment r")
```
The effect that is close to one of the anchoring effects (I2 ~75, tau ~.05) is Allowed vs. forbidden which also has r = .90.


